---
title: "STAA57 Group Project"
subtitle: "What are the dominant research areas in Ontario's automotive sector, and how do institutions specialize in different fields?"
author: "Minh Tran - 1006804914, Rayaan Syed - 1010231081"
university: "University of Toronto - Scarborough"
date: "March 15, 2025"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
# Insert all library required here
library(tidyverse)
library(readxl)
library(dplyr)
library(knitr)
library(kableExtra) 
```

```{r, include = FALSE }
# Read the main data set in as researchers and tags file as researcher_tags

researchers = read.csv("ontarioautoresearchers.csv")
researcher_tags = read_excel("ontarioautoresearcherstags.xls")
```

\pagebreak

# I. Introduction

## Research Goal

With the growing popularity of electric vehicles (EVs) and the rapid advancements in artificial intelligence (AI), particularly in the development of autonomous vehicles like Tesla, it is crucial for automotive research facilities in Ontario to focus their efforts and resources on the right areas. This will ensure a significant contribution to the academia world while assisting the government in implementing effective policies and educational institutions better educate, prepare the next generation of labors. 

This report aims to examine the automotive research areas prioritizing by Ontario's research facilities and identify specialized facilities in trending fields. It will also explore which areas are attracting the most researchers and which institutions are specializing in particular research domains.

## Dataset description

### Data source

Our dataset was collected by researchers from teh Ministry of Economic Development, Job creation and Trade and was published on the Government of Ontario's public database in 2018.

**Table 1: Data set Variables**

```{r, echo=FALSE, eval=TRUE}
glimpse(researchers)
```

Based on a summary of the data set, the `Noise..Viration.and.Harshness` variable was corrupted during the process of uploading or collecting the data and therefore will not be used in our report

```{r}
# Remove variable with errors

researchers = researchers %>% select(!Noise..Vibration.and.Harshness )

# We removed Brock University, Lakehead University, Royal Military College, and Laurentian University from our dataset as each contained less than two entries. These universities likely did not report their full research info to the dataset collectors, thus we removed it to maintain accuracy. 

researchers <- researchers %>% filter(!Institution %in% c("Brock University", "Lakehead University", "Laurentian University", "Royal Military College"))

```

Some of the main categorical variables that can help categorizing the data include

- **Researcher.Name**

- **Institution**: Name of university or research center that the researcher works at

- **Research.Areas**: Researcher general field of research

- **Tag**: Researcher's specialized field of research

In addition, the data set also contains research fields as variable (e.g. Alternative Fuels, Autonomy and AI, Vehicle Design) to indicate whether or not a researcher work fall into that field of research

# II. Data Overview Analysis

## *Descriptive Statistic*

From our dataset, we are able to extract the disparity of researchers across fields of research. This would include research fields such as "Network and Security", "Autonomy and AI", "Nanotechnology", and other vehicle-related fields.  
    
    
```{r,echo=FALSE}
# Identifying the amount of researchers across the fields of automotive design.
research_summary = researchers %>%
  pivot_longer(cols = starts_with("Tag."), names_to = "Tag_Number", values_to = "Research_Area") %>%
  filter(!is.na(Research_Area) & Research_Area != "") %>%  # Sanity check
  count(Institution, Research_Area, sort = TRUE) %>%  # Count occurrences
  pivot_wider(names_from = Research_Area, values_from = n, values_fill = 0)  # Use wide format

print(research_summary)

```
  Upon reviewing the table, the three leading researching fields are "Networks and Security", "Autonomy and AI", and "Software". The universities have reported 50+ researchers working in these ares which would indicate there is significant progress being made. On the contrast, "Crashworthiness" and "Mechatronics" are among the least studied fields with less than 10 researchers total.
  
  
  
We now check for the dominating research field for each university. The table below is generated by finding the max researchers in an institution across a research field. 
    
    
```{r,echo=FALSE}
# Identifying the dominant field for each university
    
research_summary$Institution = research_summary$Institution %>%
  str_to_title() %>%  # Brief sanity check to make sure institution strings are one. 
  str_trim()    # Ex. "Ryerson University" should be equal to "Ryerson university    " 


top_research_by_uni = research_summary %>%
  pivot_longer(cols = -Institution, names_to = "Research_Area", values_to = "Count") %>%
  group_by(Institution) %>%
  slice_max(Count, n = 1, with_ties = FALSE) %>%  # Get the top research category per institution
  select(Institution, Research_Area)

print(top_research_by_uni)

```
  From the table, the universities are associated with their highest producing research field. Several universities are mostly researching "Networks and Security" like Carleton University, University of Waterloo, University of Ontario Institute of Technology. 
  
  This information can also be crucial to understand how the fields with low researchers from the previous table are not being targeted enough by the universities.
  
  
  

We now check for the percentage breakdown of the university's dominating research field and their contribution as a whole. This is important to understand the concentration of university's in specific fields and whether some diversification can be made to the university.     
```{r, echo= FALSE}
# Identifying percentage breakdown of each university and their study on each field. 

research_summary = research_summary %>%
  mutate(Total_Research = rowSums(across(-Institution)))  # Sum all research areas per university

top_research_by_uni = research_summary %>%
  pivot_longer(cols = -c(Institution, Total_Research), names_to = "Research_Area", values_to = "Count") %>%
  mutate(Percentage = (Count / Total_Research) * 100) %>%  # Find percentage
  group_by(Institution) %>%
  slice_max(Count, n = 1, with_ties = FALSE) %>%  # Get top research area per institution
  select(Institution, Research_Area, Count, Percentage) %>% arrange(desc(Percentage)) # We want descending order for readability
top_research_by_uni


```

From the table, we see that Carleton University, York University, and University of Ottawa have 20%+ of their research contributions among their highest fields. Waterloo University and University Of Ontario Institute Of Technology have the most diversification as well as high amounts of researchers across several fields. 

## *Graph & Visualizations*

-   Bar chart: Research distribution across institution
-   Pie chart: Proportion of of different research fields
-   History, frequency of research tag

# III. Statistical Analysis

We could pick an area we think that will be most trending (AI for example) and test to see if that is true

- Confidence interval for the average number of researcher 

- Hypothesis Testing

- Bootstrapping

# IV. Predictive modeling (Regression)

Build a regression model to predict number of researcher in a field, institutions

Build a regression model to see if an institution is likely to have to a grant funding

Using cross-validation to validate the 2 models built above

# V. Summary

# Appendix
